---
title: "Test-Driving `knitr::read_chunk()` to Include External Script Chunks in an RMarkdown Document"
author: "Eric Buhle"
date: "2019-04-30 (updated `r format(Sys.time(), '%Y-%m-%d')`)"
output: 
  html_document: 
    toc: TRUE
    toc_float: TRUE
  github_document: 
    toc: TRUE
---

```{r setup, include=FALSE}
library(here)
library(knitr)
knitr::opts_chunk$set(tidy = FALSE, highlight = TRUE, comment = NA, 
                      dev = "png", dev.args = list(type = "cairo-png"), dpi = 300,
                      fig.align = "center")
```

This demo follows a helpful [blog post](http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/) by Zev Ross that shows how to import code chunks from an external script into an RMarkdown document. Also see [here](https://rpubs.com/thaufas/450838) for some more creative usage.

RMarkdown is great, but IMO it is unsuitable as an environment for code development, especially for complex analyses. The formatting and interwoven text are distracting and make re-running code a hassle; the chunk tool buttons take up screen real estate and the auto-line wrapping is the enemy of legibility; debugging is a real PITA; and I *hate* the inline presentation of figures and output when running interactively. R scripts are superior in all these respects. Until now, my strategy has been to copy-paste code chunks from a script into an RMarkdown document, a hacky approach with obvious drawbacks. Usually I end up tweaking and refining the code in the .Rmd (as opposed to doing it in the script and having to remember to copy-paste it again) and so the two documents get out of sync.

The "externalization" strategy shown here offers the best of both worlds:

* The flexibility and interactivity of R scripts for code development
* The power of RMarkdown as a format for literate programming and sharing the results of analyses

# Read chunks

First, right after the `setup` chunk, we call `knitr::read_chunk()` to ingest the code chunks from one or more scripts. (This part is crucial: we can read chunks from as many "child" scripts as we want, so long as the chunk names don't conflict!) Typically we would use `echo = FALSE` to read in the script's chunks silently. Note that the code is not executed at this point; it is just made available to be invoked in subsequent chunks. (BTW, the [RMarkdown hieroglyphics](https://yihui.org/en/2017/11/knitr-verbatim-code-chunk/) needed to produce the following verbatim RMarkdown code chunk is far more esoteric than code externalization. On the upside, it apparently makes the aforementioned chunk tool buttons vanish throughout the document.)

````
`r ''````{r read_chunk, echo=FALSE}
read_chunk(here("chunk_test1.R"))
read_chunk(here("chunk_test2.R"))
```
````

```{r read_chunk, echo=FALSE}
read_chunk(here("chunk_test1.R"))
read_chunk(here("chunk_test2.R"))
```

Let's see what's in these external scripts. This is `chunk_test1.R`:

```{r chunk_test1, echo=FALSE, warning=FALSE}
writeLines(readLines(here("chunk_test1.R")))
```

Note the special **knitr** comment syntax `## @knitr chunk_name` that demarcates a new chunk. As the regular comments (preceded by `#`) helpfully explain, the empty `## @knitr` can be used to delimit the end of a chunk followed by some lines we don't want to use in the RMarkdown (for example, verbose comments that are redundant with the body text).

Here is `chunk_test2.R`:

```{r chunk_test2, echo=FALSE, warning=FALSE}
writeLines(readLines(here("chunk_test2.R")))
```

The chunk `needs_external_input`, counterintuitively, needs external input. Specifically, `newvar` has not been assigned, so those lines as written would throw an error. We are going to assign `newvar` in an RMarkdown chunk and "pass" it to `needs_external_input`. We'll also see if the results update when we reassign `newvar` and invoke the external chunk again, and whether the behavior differs for the `print()` and `plot()` statements.

# Child script 1

## Create a data frame

The RMarkdown chunk

````
`r ''````{r variables_xy}
```
````

produces this:

```{r variables_xy}
```

## Make a plot

This time we'll use `echo = FALSE` to suppress the code.

````
`r ''````{r plot_xy, fig.height=5, fig.width=5, out.width = "40%", echo = FALSE}
```
````

```{r plot_xy, fig.height=5, fig.width=5, out.width = "40%", echo = FALSE}
```

## Fit a linear model

The next chunk shows how to fit a linear model to `y` as a function of `x`, but doesn't actually do so (`eval = FALSE`). RMarkdown:

````
`r ''````{r lm_xy, eval = FALSE}
```
````

Result:

```{r lm_xy, eval = FALSE}
```

# Child script 2

Now we'll use a chunk from the second script, entirely unrelated to the first.

## Silly `ggplot2` demo

The RMarkdown syntax

````
`r ''````{r ggplot2_demo, fig.width=7, fig.height=3, out.width="75%"}
```
````

does this, for whatever reason:

```{r ggplot2_demo, fig.width=7, fig.height=3, out.width="75%"}
```

## Chunk with inputs from RMarkdown

Can we invoke an external chunk that relies on objects defined here in the RMarkdown itself? Let's find out. First we assign a variable `newvar` with a chunk label foreshadowing that we may want to do this again.

````
`r ''````{r assign_newvar1}
newvar <- 1:10
```
````

```{r assign_newvar1}
newvar <- 1:10
```


Then we call a chunk that prints its contents and plots it with the default method.

````
`r ''````{r needs_external_input1, fig.height=5, fig.width=5, out.width = "40%"}
<<needs_external_input>>
```
````

That produces the echoed R code followed by the results of the `print()` and `plot()` calls, respectively:

```{r needs_external_input1, fig.height=5, fig.width=5, out.width = "40%"}
<<needs_external_input>>
```

And what if we modify `newvar` and then call the summary chunk again? The output is:

```{r assign_newvar2}
newvar <- runif(10,0,1)
```
```{r needs_external_input2, fig.height=5, fig.width=5, out.width = "40%"}
<<needs_external_input>>
```

This works, but the plot call **only** works if we use the `<<chunk_name>>` syntax as opposed to the `{r chunk_name}` syntax. Both are described in the Zev Ross post linked above. If we use the latter, the plots are identical and both use the last-defined value of `newvar`. I don't understand why, but it must be something about how the output from the .Rmd code chunks are knitted together. The `<<chunk_name>>` syntax produces an "unexpected token" warning next to the chunk in the editor window, which is annoying, but it gives the correct plots.

