---
title: "Test-Driving `knitr::read_chunk()` to Include External Script Chunks in an RMarkdown Document"
author: "Eric Buhle"
date: "2019-04-30 (updated `r format(Sys.time(), '%Y-%m-%d')`)"
output: 
  html_document: default
  github_document: default
---

```{r setup, include=FALSE}
library(here)
library(knitr)
knitr::opts_chunk$set(tidy = FALSE, highlight = TRUE, comment = NA, 
                      dev = "png", dev.args = list(type = "cairo-png"), dpi = 300,
                      fig.align = "center")
```

This example follows a helpful [blog post](http://zevross.com/blog/2014/07/09/making-use-of-external-r-code-in-knitr-and-r-markdown/) by Zev Ross that shows how to import code chunks from an external script into an RMarkdown document. Also see [here](https://rpubs.com/thaufas/450838) for some more creative usage.

RMarkdown is great, but IMO it is unsuitable as an environment for code development, especially for complex analyses. The formatting and interwoven text are distracting and make re-running code a hassle; the chunk tool buttons take up screen real estate and the auto-line wrapping is the enemy of legibility; debugging is a real PITA; and I *hate* the inline presentation of figures and output when running interactively. R scripts are superior in all these respects. Until now, my strategy has been to copy-paste code chunks from a script into an RMarkdown document, a hacky approach with obvious drawbacks. Usually I end up tweaking and refining the code in the .Rmd (as opposed to doing it in the script and having to remember to copy-paste it again) and so the two documents get out of sync.

The "externalization" strategy shown here offers the best of both worlds:

* The flexibility and interactivity of R scripts for code development
* The power of RMarkdown as a format for literate programming and sharing the results of analyses

# Read chunks

First, right after the `setup` chunk, we call `knitr::read_chunk()` to ingest the code chunks from one or more scripts. (This part is crucial: we can read chunks from as many "child" scripts as we want, so long as the chunk names don't conflict!) Ordinarily we would use `echo = FALSE` to read in the script's chunks silently. Note that the code is not executed at this point; it is just made available to be invoked in subsequent chunks.

```{r read_chunks}
read_chunk(here("chunk_test1.R"))
read_chunk(here("chunk_test2.R"))
```

Let's see what's in these external scripts. This is `chunk_test1.R`:

```{r chunk_test1, echo=FALSE, warning=FALSE}
writeLines(readLines(here("chunk_test1.R")))
```

Note the special **knitr** comment syntax `## @knitr chunk_name` that demarcates a new chunk. As the regular comments (preceded by `#`) helpfully explain, the empty `## @knitr` can be used to delimit the end of a chunk followed by some lines we don't want to use in the RMarkdown (for example, verbose comments that are redundant with the body text).

Here is `chunk_test2.R`:

```{r chunk_test2, echo=FALSE, warning=FALSE}
writeLines(readLines(here("chunk_test2.R")))
```

The chunk `needs_external_input`, counterintuitively, needs external input. Specifically, `newvar` has not been assigned, so those lines as written would throw an error. We are going to assign `newvar` in an RMarkdown chunk and "pass" it to `needs_external_input`. We'll also see if the results update when we reassign `newvar` and invoke the external chunk again, and whether the behavior differs for the `print()` and `plot()` statements.

Now let's use the **knitr** chunks we've ingested from these two external scripts.

# Script 1

## Run the `variables_xy` chunk and use the variables it creates
```{r variables_xy}
```

## Run the `plot_xy` chunk and create the plot

This time we'll use `echo = FALSE` to suppress the code.

```{r plot_xy, fig.height=5, fig.width=5, out.width = "40%", echo = FALSE}
```

## Fit linear model

The next chunk shows how to fit a linear model to `y` as a function of `x`, but doesn't actually do so (`eval = FALSE`).

```{r lm_xy, eval = FALSE}
```

# Script 2

Now we'll use a chunk from the second script, entirely unrelated to the first.

## Silly `ggplot2` demo

```{r ggplot2_demo, fig.width=7, fig.height=3, out.width="75%"}
```

## Chunk that requires objects defined in RMarkdown

Can we invoke an external chunk that relies on objects defined here in the RMarkdown itself? Let's find out. First we assign a variable `newvar`, and then we call a chunk that prints its contents and plots it with the default method.

```{r assign_newvar1}
newvar <- 1:10
```
```{r needs_external_input1, fig.height=5, fig.width=5, out.width = "40%"}
<<needs_external_input>>
```

And what if we modify `newvar` and then call the summary chunk again?

```{r assign_newvar2}
newvar <- runif(10,0,1)
```
```{r needs_external_input2, fig.height=5, fig.width=5, out.width = "40%"}
<<needs_external_input>>
```

This works, but the plot call **only** works if we use the `<<chunk_name>>` syntax as opposed to the `{r chunk_name}` syntax. Both are described in the Zev Ross post linked above. If we use the latter, the plots are identical and both use the last-defined value of `newvar`. I don't understand why, but it must be something about how the output from the .Rmd code chunks are knitted together. The `<<chunk_name>>` syntax produces an "unexpected token" warning next to the chunk in the editor window, which is annoying, but it gives the correct plots.

